### 2020-02-02 abstract_file.tex
minor changes

In many applications, one must work with models that have been trained by someone else.
For such \emph{pretrained models}, one typically does not have access to training data or test data, and one does not know the details of how the models were built, the specifics of the data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized, etc.
Here, we present and evaluate quality metrics for pretrained neural network models at scale.
%%Our metrics are drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as Heavy-Tailed Random Matrix Theory (HT-RMT), as used in the recently-developed Theory of Heavy-Tailed Self Regularization (e.g., fitted power law metrics used to characterize the degree of strong correlations in trained models).
%%The most promising are metrics drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as metrics (e.g., fitted power law metrics used to characterize the degree of strong correlations in a system) derived from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
The most promising are norm-based metrics (such as those used to provide capacity control in traditional statistical learning theory) and metrics based on fitting eigenvalue distributions to truncated power law (PL) distributions (which derive from statistical mechanics, in particular the recently-developed Theory of Heavy-Tailed Self Regularization, and which can be used to characterize the degree of strong correlations in a system). 
%
Using the publicly-available \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision and natural language processing.
We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can be used to discriminate between ``good-better-best.'' 
On the other hand, for models that may not be well-trained (which, arguably is the point of needing metrics to evaluate the quality of pretrained models), i.e., when we want to distinguish ``good-vs-bad,'' norm-based metrics can qualitatively fail. 
We also find that PL-based metrics do much better, quantitatively better at discriminating good-better-best and qualitatively better at discriminating good-vs-bad.
PL-based metrics can also be used to characterize fine-scale properties of models, e.g., understanding layer-wise \emph{correlation flow}, and evaluate post-training modifications such as model distillation and model improvement.

>>>>

In many practical AI applications, one works with deep neural network (DNN) models trained by someone else.
For such \emph{pretrained models}, one typically does not have access to training data or test data, and, moreover,  does not know the details of how the models were built, the specifics of the data that were used to train the model, what was the loss function or hyperparameter values, how precisely the model was regularized, etc.
Here, we present and evaluate quality metrics for pretrained neural network models at scale.
%%Our metrics are drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as Heavy-Tailed Random Matrix Theory (HT-RMT), as used in the recently-developed Theory of Heavy-Tailed Self Regularization (e.g., fitted power law metrics used to characterize the degree of strong correlations in trained models).
%%The most promising are metrics drawn from traditional statistical learning theory (e.g., norm-based capacity control metrics) as well as metrics (e.g., fitted power law metrics used to characterize the degree of strong correlations in a system) derived from the recently-developed Theory of Heavy-Tailed Self Regularization (HT-SR).
The most promising are norm-based metrics (such as those used to provide capacity control in traditional statistical learning theory) and metrics based on fitting eigenvalue distributions to truncated power law (PL) distributions (which derive from statistical mechanics, in particular the recently-developed Theory of Heavy-Tailed Self Regularization, and characterize the strength of correlations in a DNN). 
%
Using the publicly-available \emph{WeightWatcher} tool, we analyze hundreds of publicly-available pretrained models, including older and current state-of-the-art models in computer vision (ZCV) and natural language processing (NLP)
We find that norm-based metrics do a reasonably good job at predicting quality trends in well-trained models, i.e., they can discriminate amoing ``good-better-best''  models.
On the other hand, for poorly trained models--which, arguably why we need quality metrics-- we want to distinguish which and when norm-based metrics can distiungush ''good-vs-bad'' models.
We also find that PL-based metrics do much better, quantitatively better, at discriminating aming ''good-better-best'' models, and qualitatively better at discriminating ''good-vs-bad'' models.
PL-based metrics can also be used to characterize fine-scale properties of models, e.g., understanding layer-wise \emph{correlation flow}, and evaluate post-training modifications such as model distillation, increasing the data set size, and other model improvements.

### 2020-02-02 background.tex
pass over most sections, more to do on sec 5 and 6
minor changes

Most theory for DNNs is applied to small toy models and assumes access to data.
There is very little work asking how to predict, in a theoretically-principled manner, the quality of large-scale state-of-the-art NNs, and how to do so without access to training data or testing data or details of the training protocol, etc.
Our work is, however, loosely related to several other lines of work.

\paragraph{Statistical mechanics theory for NNs.}

Statistical mechanics ideas have long had influence on NN theory and practice~\cite{EB01_BOOK, MM17_TR, BKPx20}; and 
our best-performing metrics (those using fitted PL exponents) are based on statistical mechanics~\cite{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM19_KDD, MM20_SDM}, in particular the recently-developed \emph{Theory of Heavy Tailed Self Regularization (HT-SR)}~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.  
However, we emphasize that the way in which we \emph{use} statistical mechanics theory is quite different than the way it is more commonly formulated.
Several very good overviews of the more common approach are available~\cite{EB01_BOOK, BKPx20}.
We will \emph{use} statistical mechanics theory more as it is used in quantitative finance.
Thus, much more relevant for our metholodogical approach is older work of Bouchaud, Potters, Sornette, and coworkers~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} on the statisical mechanics of heavy tailed and strongly correlated systems.

@@

There is also a large body of work on using norm-based metrics to bound generalization error~\cite{NTS15, BFT17_TR, LMBx18_TR}.
In this area, theoretical work aims to prove generalization bounds, and applied work uses these norms as a regularization function to improve training.
While we do find that norms provide relatively good quality metrics, at least for distinguishing good-better-best among well-trained models, we are not interested in proving generalization bounds or improving training.

\paragraph{Practical problems poorly addressed by theory.}
There are many very practical problems in ML that are poorly addressed by existing theory and that either motivated our work or should be addressable by our techniques.
Here are several.
\begin{itemize}
\item
\textbf{Simplicity, or lack of,  accuracy metrics.}
\michael{XXX.  Put in two or at most three sentences.}
\item
\textbf{Information leakage in the production pipeline.}
\michael{XXX.  Put in two or at most three sentences.}
\item
\textbf{Cost of acquring labeled data.}
\michael{XXX.  Put in two or at most three sentences.}
\end{itemize}
Importantly, there are many examples in ML where (as a practical matter) there is no reliable notion of accuracy, e.g., when generating fake text or realiztic chatbots, developing self driving cars, or just clusterting user profiles.
\footnote{For example, current chatbots use perplexity as a proxy for passing a Turing test.}
%, and when distilling a reliable model in some way to obtain comparable training/test quality but that damages the model in some other subtle way. 
%\michael{That last example is awkward.  It would be good to have a better example and plant seeds for model distillation elsewhere.}

>>>>

Most theory for DNNs is applied to small toy models and assumes access to data.
There is very little work asking how to predict, in a theoretically-principled manner, the quality of large-scale state-of-the-art DNNs, and how to do so without access to training data or testing data or details of the training protocol, etc.
Our work is, however, loosely related to several other lines of work.

\paragraph{Statistical mechanics theory for DNNs.}

Statistical mechanics ideas have long had influence on DNN theory and practice~\cite{EB01_BOOK, MM17_TR, BKPx20}; and 
our best-performing metrics (those using fitted PL exponents) are based on statistical mechanics~\cite{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM19_KDD, MM20_SDM}, in particular the recently-developed \emph{Theory of Heavy Tailed Self Regularization (HT-SR)}~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.  
However, we emphasize that the way in which we \emph{use} statistical mechanics theory is quite different than the way it is more commonly formulated.
Several very good overviews of the more common approach are available~\cite{EB01_BOOK, BKPx20}.
We will \emph{use} statistical mechanics in a broader sense, drawing upon results from quantitative finance and random matrix theory (RMT).
Thus, much more relevant for our metholodogical approach is older work of Bouchaud, Potters, Sornette, and coworkers~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} on the statisical mechanics of heavy tailed and strongly correlated systems\charles{, and current work (in progress)\cite{blog}}

@@

There is also a large body of work on using norm-based metrics to bound generalization error~\cite{NTS15, BFT17_TR, LMBx18_TR}.
In this area, theoretical work aims to prove generalization bounds, and applied work uses these norms as a regularization function to improve training.
While we do find that norms provide relatively good quality metrics, at least for distinguishing good-better-best among well-trained models, we are not interested in proving generalization bounds \nred{or developing new regularizers.}

\paragraph{Practical problems need a practical theory.}
\nred{Exiting ML theory does not address many practical AI problems that our techiques directly target.}
%There are many very practical problems in ML that are poorly addressed by existing theory and that either motivated our work or should be addressable by our techniques.
Here are several.
\begin{itemize}[leftmargin=*]
\item
\textbf{Simplicity, or lack of,  accuracy metrics.}
\michael{XXX.  Put in two or at most three sentences.}
\item
\textbf{Information leakage in the production pipeline.}
\michael{XXX.  Put in two or at most three sentences.}
\item
\textbf{Cost of acquring labeled data.}
\michael{XXX.  Put in two or at most three sentences.}
\end{itemize}

\charles{see comments in tex file}
% we adress this above
%Importantly, there are many examples in ML where (as a practical matter) there is no reliable notion of accuracy, e.g., when generating fake text or realiztic chatbots, developing self driving cars, or just clusterting user profiles.
%\footnote{For example, current chatbots use perplexity as a proxy for passing a Turing test.}
%, and when distilling a reliable model in some way to obtain comparable training/test quality but that damages the model in some other subtle way. 
%\michael{That last example is awkward.  It would be good to have a better example and plant seeds for model distillation elsewhere.}

### 2020-02-02 conclusion.tex
minor changes

We conclude with a few thoughts on what a \emph{practical theory} of NNs should look like.
%
We distinguish between what we will call a
\emph{phenomenological theory}
(that describes empirical relationship of phenomena to each other, in a way which is consistent with fundamental theory, but is not directly derived from that theory)
and what can be called a 
\emph{first principles theory} 
(that is applicable to toy models, but that does not scale up to realistic systems if one includes realistic apsects of realistic systems).
%
For most complex highly-engineered systems (aside from complex AI/ML systems), one \emph{uses} phenomenological theory rather than first principles theory. 
(One does not try to solve the Schr\"odinger equation if one is interested in building a bridge or flying an airplane.)
%
Our results, which are based on our \emph{use} of sophisticated statistical mechanics theory to solve an important practical NN problems, suggests that this approach should be of interest more generally for those interested in developing a practical NN theory.

>>>>

We conclude with a few thoughts on what a \emph{practical theory} of DNNs should look like.
%
We distinguish between what we will call a
\emph{phenomenological theory}
(that describes empirical relationship of phenomena to each other, in a way which is consistent with fundamental theory, but is not directly derived from that theory)
and what can be called a 
\emph{first principles theory} 
(that is applicable to toy models, but that does not scale up to realistic systems if one includes realistic apsects of realistic systems).
%
For most complex highly-engineered systems (aside from complex AI/ML systems), one \emph{uses} phenomenological theory rather than first principles theory. 
(One does not try to solve the Schr\"odinger equation if one is interested in building a bridge or flying an airplane.)
%
Our results, which are based on our \emph{use} of sophisticated statistical mechanics theory to solve an important practical DNN problems, suggests that this approach should be of interest more generally for those interested in developing a practical DNN theory.

### 2020-02-02 cv_models.tex
some misc stuff
minor changes

@@

\michael{XXX.  WORK ON THIS PAR AFTER DO NLP EXAMPLES.}
Our interpretation of this is the following.
The pre-trained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than any of the VGG/ResNet/DenseNet models we condisered previously.
Norms are often used as regularizers, and thys they are a natural metric to consider, e.g., during training or distillation.
The PL $\alpha$ captures more subtle correlations.
%
The reason for this is that the \texttt{distiller} Group Regularization technique has the unusual effect of increasing the norms of the $\mathbf{W}$ feature maps for at least two of the Conv2D layers.
\michael{This impedance shuffling or whatever it is called messes up norms, but the strong correlations are still preserved, as seen in the $\hat{\alpha}$ metric, and the model quality is good.  CLARIFY.}
\michael{XXX.  INTERPRET IN TERMS OF XXX GOOD VERSUS BAD, AS OPPOSED TO GOOD BETTER BEST.}

>>>>

%MM% \begin{figure}
%MM% \centering
%MM% \begin{minipage}[b]{.22\textwidth}
%MM% %<Code for the first figure>
%MM% XXX
%MM% \caption{Caption}\label{label-a}
%MM% \end{minipage}\qquad
%MM% \begin{minipage}[b]{.22\textwidth}
%MM% %<Code for the second figure>
%MM% XXX
%MM% \caption{Caption}\label{label-b}
%MM% \end{minipage}
%MM% \end{figure}

@@

\michael{XXX.  WORK ON THIS PAR AFTER DO NLP EXAMPLES.}
Our interpretation of this is the following.
The pre-trained models in the \texttt{distiller} package have passed some quality metric, but they are much less well trodden than any of the VGG/ResNet/DenseNet models we condisered previously.
\charles{MOvE THIS SOMEWHERE ELSE AND CLEAN UP THIS SECTION Notice while norms make good regularizers for a single model, there is no reason \emph{a priori} 
to expect them correlate so well with the test accuracies across different models.
However, we do expect the PL $\alpha$ to do so because it effectively measures the amount of correlation in the model}
%
The reason for this is that the \texttt{distiller} Group Regularization technique 
%%has the unusual effect of increasing
\nred{spuriously increases}
 the norms of the $\mathbf{W}$ pre-activation maps for at least two of the Conv2D layers.
\michael{This impedance shuffling or whatever it is called messes up norms, but the strong correlations are still preserved, as seen in the $\hat{\alpha}$ metric, and the model quality is good.  CLARIFY.}
\michael{XXX.  INTERPRET IN TERMS OF XXX GOOD VERSUS BAD, AS OPPOSED TO GOOD BETTER BEST.}

### 2020-02-02 introduction.tex
minor changes

A common problem in machine learning (ML) 
%and artificial intelligence (AI) 
is to evaluate the quality of a given model.
A popular way to accomplish this
%, in particular in academic environments, 
is to train a model and then evaluate its training/testing error.
There are many problems with this approach.
Well-known problems with just examining training/testing curves include that 
they give very limited insight into the overall properties of the model, 
they do not take into account the (often large human and CPU/GPU) time for hyperparameter fiddling,
they typically do not correlate with other properties of interest such as robustness or fairness or interpretability, 
and so on.
A less well-known problem, but one that is increasingly important (in particular in industrial-scale ML---where the \emph{users} of models are not the \emph{developers} of the models) is that one may have access to neither the training data nor the testing data.
Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and be told to use it.

@@

Moreover, in industry, one faces unique practical problems such as: do we have enough data for this model? 
Methods that are developed and evaluated on any well-defined publicly-available coprus of data, no matter how large or diverse or interesting, are clearly not going to be well-suited to address problems such as this.
It is of great practical interest to have metrics to evaluate the quality of a ML model---in the absence of training/testing data and without any detailed knowledge of the training/testing process.  
We seek a practical theory for pretrained models which can predict how, when, and why such models can be expected to perform well or~poorly.

In this paper, we present and evaluate quality metrics for pretrained neural network (NN) and deep neural network (DNN) models at scale.
To do so, we consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
%
By now, there are many such state-of-the-art models that are publicly-available, e.g., 
there are now hundreds of pretrained models in CV ($\ge 500$) and NLP ($\approx 100$).%
\footnote{When we began this work in 2018, there were fewer than tens of such models; now in 2020, there are hundreds of such models; and we expect that in a year or two there will be an order of magnitude or more of such models.}
These provide a large corpus of models that by some community standard are state-of-the-art.%
\footnote{Clearly, there is a selection bias or survivorship bias here---people tend not to make publicly-available their poorly-performing models---but these models are things in the world that (like social networks or the internet) can be analyzed for their properties.}
Importantly, all of these models have been trained by someone else and have been viewed to be of sufficient interest/quality to be made publicly-available; and, for all of these models, we have no access to training data or testing data, and we have no knowledge of the training/testing protocols.

The \emph{quality metrics} we consider are based on the spectral properties of the layer weight matrics.
In particular, they are based on the norms of weight matrices (such norms have been used in traditional statistical learning theory to bound capacity and construct regularizers) and parameters of power law (PL) fits of the eigenvalues of weight matrices (such PL fits are based on statistical mechanics approaches to NNs).
Note that, while we use traditional norm-based and PL-based metrics, our goals are different.
In particular, unlike more common ML approaches, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., hyperparmeter optimization).% 
\footnote{One could of course use these techniques to improve training, and we have been asked about that, but we are not interested in that here. Our main goal here is to use these techniques to evaluate properties of state-of-the-art pretrained NN models.}
Instead, we want to examine different models across common architecture series, and we want to compare models between different architectures themselves, and in both cases \emph{we aim to predict trends in the quality of pre-trained models without access to training or testing data}.

@@

\noindent
We emphasize that our goal is a practical theory to predict trends in the quality of state-of-the-art NN models, i.e., not to make a statement about every publicly-available model.
We have examined hundreds of models, and are exceptions to the general trends we state.
Several of the most interesting are described below.

@@

All of our computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package};
and, in order to make our results fully reproducible, we provide Jupyter notebooks in the github repo for this paper~\cite{repo},
with fully reproducible Jupyter notebooks.%
\footnote{\url{https://github.com/CalculatedContent/kdd2020} \michael{DIFFERENT NAME?}}
\michael{LINK TO: Google Colab, IS THIS THAT REPO.}
\michael{FIX THIS PAR.}

>>>>

A common problem in machine learning (ML) 
%and artificial intelligence (AI) 
is to evaluate the quality of a given model.
A popular way to accomplish this
%, in particular in academic environments, 
is to train a model and then evaluate its training/testing error--but
there are many problems with this approach.
The training/testing curves give very limited insight into the overall properties of the model, 
they do not take into account the (often large human and CPU/GPU) time for hyperparameter fiddling,
they typically do not correlate with other properties of interest such as robustness or fairness or interpretability, 
and so on.
A less well-known problem, but one that is increasingly important (in particular in industrial-scale 
artificial intelligence (AI), arises when the model \emph{user} is not the model \emph{developer}.
here, one may not have access to either the training data nor the testing data.
\nred{Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and need to use it "as is'', or fine tune it and/or compress it.}

@@

Moreover, for many large scale, practical applications, one can not define a true test metric.  For example, NLP models that generate fake text or conversational chatbots may use a proxy, like perplexity, as a test metric, but, in the end, really require human evaluation.  Or, say, models that cluster user profiles, which are widely used in marketing, are unsupervised and have no labels for comparison.  \charles{more here, clean up?}

Most importantly, in industry, one faces unique practical problems such as: do we have enough data for this model? 
\nred{Indeed, high quality, labeled data can be very expensive to acquire, and this cost can make or break a project.}
Methods that are developed and evaluated on any well-defined publicly-available coprus of data, no matter how large or diverse or interesting, are clearly not going to be well-suited to address problems such as this.
It is of great practical interest to have metrics to evaluate the quality of a trained model---in the absence of training/testing data and without any detailed knowledge of the training/testing process.  
We seek a practical theory for pretrained models which can predict how, when, and why such models can be expected to perform well or~poorly.

In this paper, we present and evaluate quality metrics for pretrained deep neural network (DNN) models, and at scale.
To do so, we consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
%
By now, there are many such state-of-the-art models that are publicly-available, e.g., 
there are now hundreds of pretrained models in CV ($\ge 500$) and NLP ($\approx 100$).%
\footnote{When we began this work in 2018, there were fewer than tens of such models; now in 2020, there are hundreds of such models; and we expect that in a year or two there will be an order of magnitude or more of such models.}
These provide a large corpus of models that by some community standard are state-of-the-art.%
\footnote{Clearly, there is a selection bias or survivorship bias here---people tend not to make publicly-available their poorly-performing models---but these models are things in the world that (like social networks or the internet) can be analyzed for their properties.}
Importantly, all of these models have been trained by someone else and have been viewed to be of sufficient interest/quality to be made publicly-available; and, for all of these models, we have no access to training data or testing data, and we have no knowledge of the training/testing protocols.

The \emph{quality metrics} we consider are based on the spectral properties of the layer weight matrics.
In particular, they are based on the norms of weight matrices (such norms have been used in traditional statistical learning theory to bound capacity and construct regularizers) and parameters of power law (PL) fits of the eigenvalues of weight matrices (such PL fits are based on statistical mechanics approaches to DNNs).
Note that, while we use traditional norm-based and PL-based metrics, our goals are different.
In particular, unlike more common ML approaches, \emph{we do not seek a bound on the generalization} (e.g., by evaluating training/test error during training), \emph{we do not seek a new regularizer}, and \emph{we do not aim to evaluate a single model} (e.g., hyperparmeter optimization).% 
\footnote{One could of course use these techniques to improve training, and we have been asked about that, but we are not interested in that here. Our main goal here is to use these techniques to evaluate properties of state-of-the-art pretrained DNN models.}
Instead, we want to examine different models across common architecture series, and we want to compare models between different architectures themselves, and in both cases \emph{we aim to predict trends in the quality of pre-trained models without access to training or testing data}.

@@

\noindent
We emphasize that our goal is a practical theory to predict trends in the quality of state-of-the-art DNN models, i.e., not to make a statement about every publicly-available model.
\nred{We have examined hundreds of models, we indentify the general trends, and highlight interesting exceptions.}
%%Several of the most interesting are described below.

@@

All of our computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package};
and, in order to make our results fully reproducible, we provide Jupyter notebooks in the github repo for this paper~\cite{repo},
and with fully reproducible Google colab notebooks for the larger calculations.%
\footnote{\url{https://github.com/CalculatedContent/kdd2020} \michael{DIFFERENT NAME?}\charles{suggestions?}}
\michael{LINK TO: Google Colab, IS THIS THAT REPO.}
\michael{FIX THIS PAR.}

### 2020-02-02 methods.tex
minor changes

Each DNN layer contains one or more layer 2D weight matrices $\mathbf{W}_{l}$, and/or 2D feature maps $\mathbf{W}_{i,l}$ extracted from 2D Convolutional layers. 
(For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below; and we let $\mathbf{W}$ be a generic $N\times M$ weight matrix, with $N\ge M$.) 
%
We have examined a large number of possible quality metrics.
The best performing metrics (recall that we can only consider metrics that do not use training/test data) depend on the norm and/or spectral properties of weight matrices, $\mathbf{W}$.%
\footnote{We do not use intra-layer information from the models in our quality metrics, but as we will describe our metrics can be used to learn about that.}
Consider the following.
\begin{itemize}
\item 
Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\sum_{i,j}w^{2}_{i,j} = \sum_{i=1}^{M} \lambda_{i}^{2}$
\item 
Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$
\michael{Should this be $\sqrt{\lambda_{max}}$?}
\item 
Weighted Alpha Metric: $\hat{\alpha}=\alpha\log\lambda_{max}$
\item 
$\alpha$-Norm (or $\alpha$-Shatten Norm): $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda_{i}^{\alpha}$
\end{itemize}
The first two quantities are well-known in ML.
The last two deserve special mention.
For all these quantities, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the \emph{Empirical Correlation Matrix},
$ %% $$
\mathbf{X}=\mathbf{W}^{T}\mathbf{W} ,
$ %% $$
and so $\lambda_{max}$ is the maximum eigenvalue of $\mathbf{X}$. 
(These eigenvalues are the squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, i.e., $\lambda_{i}=\sigma^{2}_{i}$.)
For the last two quantities, $\alpha$ is the PL exponent that arises in the recently-developed HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
Operationally, $\alpha$ is determined by using the publicly-available \emph{WeightWatcher} tool~\cite{weightwatcher_package} to fit the Empirical Spectral Density (ESD) of $\mathbf{X}$, i.e., a histogram of the eigenvalues, call it $\rho(\lambda)$, to a truncated PL, 
\begin{equation}
\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\le\lambda_{max}  ,
\end{equation}
where $\lambda_{max}$ \emph{is} the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
Each of these quantities is defined for a given layer $\mathbf{W}$ matrix.

@@

\paragraph{Normalization of Empirical Matrices.}  
Normalization is an important, if underappreciated, practical issue.
Importantly, the normalization of weight matrices does \emph{not} affect the Power Law fits because the Heavy Tailed exponent $\alpha$ (as well as other metrics such as the Stable Rank and MP Soft Rank~\cite{MM18_TR,MM19_HTSR_ICML}) is scale-invariant.
Norm-based metrics, however, do depend strongly on the scale of the weight matrix.
Typically, to apply RMT, we would usually define Correlation Matrix with $1/N$ normalization and assume that the variance of $\mathbf{X}$ is either unity or a known constant.% 
\footnote{For Heavy Tailed theorems, one typically needs a normalization such as $1/N^{1-\alpha}$.}
Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with the variance already normalized to $\sqrt{1/N}$, or some variant of this, e.g., the Glorot/Xavier normalization~\cite{GloRot}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers.
We do not have conrol over the final empirical normalization of these models; and we do \emph{not} normalize (or renormalize) the Empirical Correlation Matrices, i.e., we use them as-is.
The only exception to this is that we do rescale the Conv2D pre-Activation Maps $\mathbf{W}_{i,L}$ by $k/\sqrt{2}$ so that they are on the same scale as the Linear layers.

>>>>

Each DNN layer contains one or more layer 2D weight matrices $\mathbf{W}_{l}$, and/or 2D feature maps $\mathbf{W}_{i,l}$ extracted from 2D Convolutional layers. 
(For notational convenience, we may drop the $i$ and/or $i,l$ subscripts below; and we let $\mathbf{W}$ be a generic $N\times M$ weight matrix, with $N\ge M$.) 
%
We have examined a large number of possible quality metrics.
The best performing metrics (recall that we can only consider metrics that do not use training/test data) depend on the norm and/or spectral properties of weight matrices, $\mathbf{W}$.%
\footnote{We do not use intra-layer information from the models in our quality metrics, but as we will describe our metrics can be used to learn about that.}
Consider the following.
\charles{Note I write X for the $\alpha$-norm and not W.  This is because it would be $2\alpha$.  I know its a bit confusing.
We have to be careful here with the Frobenius norm also and check that we are not off by a factor of 2}
\begin{itemize}
\item 
Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\sum_{i,j}w^{2}_{i,j} = \sum_{i=1}^{M} \lambda_{i}^{2}$
\item 
Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}=\lambda_{max}$
\michael{Should this be $\sqrt{\lambda_{max}}$?}
\item 
Weighted Alpha Metric: $\hat{\alpha}=\alpha\log\lambda_{max}$
\item 
$\alpha$-Norm (or $\alpha$-Shatten Norm): $\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}=\sum_{i=1}^{M}\lambda_{i}^{\alpha}$
\end{itemize}
The first two quantities are well-known in ML.
The last two deserve special mention.
For all these quantities, $\lambda_{i}$ is the $i^{th}$ eigenvalue of the \emph{Empirical Correlation Matrix},
$ %% $$
\mathbf{X}=\mathbf{W}^{T}\mathbf{W} ,
$ %% $$
and so $\lambda_{max}$ is the maximum eigenvalue of $\mathbf{X}$. 
(These eigenvalues are the squares of the singular values $\sigma_{i}$ of $\mathbf{W}$, i.e., $\lambda_{i}=\sigma^{2}_{i}$.)
For the last two quantities, $\alpha$ is the PL exponent that arises in the recently-developed HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
Operationally, $\alpha$ is determined by using the publicly-available \emph{WeightWatcher} tool~\cite{weightwatcher_package} to fit the Empirical Spectral Density (ESD) of $\mathbf{X}$, i.e., a histogram of the eigenvalues, call it $\rho(\lambda)$, to a truncated PL, 
\begin{equation}
\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\le\lambda_{max}  ,
\end{equation}
where $\lambda_{max}$ \emph{is} the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
Each of these quantities is defined for a given layer $\mathbf{W}$ matrix.

@@

\paragraph{Normalization of Empirical Matrices.}  
Normalization is an important, if underappreciated, practical issue.
Importantly, the normalization of weight matrices does \emph{not} affect the Power Law fits because the Heavy Tailed exponent $\alpha$ (as well as other metrics such as the Stable Rank and MP Soft Rank~\cite{MM18_TR,MM19_HTSR_ICML}) is scale-invariant.
Norm-based metrics, however, do depend strongly on the scale of the weight matrix--\nred{that's the point.}
\nred{Indeed, early theoretical work by Bartlett suggests that the test accuracy depends strongly on the ``total size'' of the weight matrics.}
Typically, to apply RMT, we would usually define Correlation Matrix with $1/N$ normalization and assume that the variance of $\mathbf{X}$ is either unity or a known constant.% 
\footnote{For Heavy Tailed theorems, one typically needs a normalization such as \nred{$1/N^{\alpha-1}$. check this}}
Pretrained DNNs are typically initialized with random weight matrices $\mathbf{W}_{0}$, with the variance already normalized to $\sqrt{1/N}$, or some variant of this, e.g., the Glorot/Xavier normalization~\cite{GloRot}, or a $\sqrt{2/Nk^2}$ normalization for Convolutional 2D Layers.
We do not have conrol over the final empirical normalization of these models; and we do \emph{not} normalize (or renormalize) the Empirical Correlation Matrices, i.e., we use them as-is.
The only exception to this is that we do rescale the Conv2D pre-Activation Maps $\mathbf{W}_{i,L}$ by $k/\sqrt{2}$ so that they are on the same scale as the Linear layers.
